Inferring probabilistic stellar rotation periods using Gaussian processes

### 2017-02-16 documents/ms.tex
adding some words to the conclusion

We also compare our results to those of the Tel Aviv team in
\citet{Aigrain2015} who use the ACF method described in \citet{Mcquillan2013}.
Although we follow the same overall ACF method in this paper, the exact
implementation of \citet{Mcquillan2013} code is slightly different and
therefore produces different results.
The MRD of their results relative to the true rotation periods is
\percenttelavivMAD\% which is smaller than the results produced by the GP
method: their results are more precise.
However, they produce a larger number of large outliers --- they predict
periods that are significantly different from the true period in more cases
than the GP method.
This is revealed by comparing the root-mean-squared (RMS) of the two sets of
results, which is more sensitive to outliers: the RMS of the Tel Aviv results
is 25 days and the RMS of GP results is 7 days.
The Tel Aviv ACF results are therefore more precise but are catastrophically
wrong in a larger number of cases.
% This can be understood as a consequence of the ACF method's sensitivity to
% noise.
% The ACF method will predict a period according to the positions of the highest
% peaks in the ACF which, on occasion, are produced by noise in the data rather
% than the periodic signal.
% On the other hand, the priors on the GP parameters ensure that a `reasonable'
% rotation period is always recovered --- it is therefore less sensitive to
% noise and rarely infers periods that are .

@@

We have attempted to recover the rotation periods of \naigrain\ simulated
\kepler-like light curves for solid-body rotators \citep{Aigrain2015} using
three different methods: a new Gaussian process method,
an authocorrelation function method, and a Lomb-Scargle periodogram method.
We demonstrate that the GP method produces the most accurate rotation periods
of the three techniques, providing a large improvement over the LS periodogram
method and a moderate one over the ACF method.
We also find that the standard version of the ACF method, most commonly
implemented in the literature, often slightly under-predicts the rotation
period due to a subtle feature of its peak detection algorithm.
In addition, we measure the rotation periods of \nkoimcq\ \kepler\ objects of
interest using the GP method, and find that these results compare well
to those measured previously by \citet{Mcquillan2013}.
The good agreement between these two sets of results demonstrates that the GP
method works well on real \kepler\ data.

>>>>

We also compare our results to those of the Tel Aviv team in
\citet{Aigrain2015} who use the ACF method described in \citet{Mcquillan2013}.
Although we follow the same general ACF method in this paper, the exact
implementation of the \citet{Mcquillan2013} code is slightly different and
produces different results.
The MRD of their periods, relative to the true rotation periods, is
\percenttelavivMAD\% --- smaller than the results produced by the GP
method.
However, their results contain a larger number of extreme outliers --- they
predict
periods that are significantly different from the true period in more cases
than the GP method.
This is revealed by comparing the root-mean-squared (RMS) of the two sets of
results since the RMS is more sensitive to outliers.
The RMS of the Tel Aviv method results is \telavivRMS\ days and the RMS of GP
method results is \gpRMS\ days.
% This can be understood as a consequence of the ACF method's sensitivity to
% noise.
% The ACF method will predict a period according to the positions of the highest
% peaks in the ACF which, on occasion, are produced by noise in the data rather
% than the periodic signal.
% On the other hand, the priors on the GP parameters ensure that a `reasonable'
% rotation period is always recovered --- it is therefore less sensitive to
% noise and rarely infers periods that are .

@@

We have attempted to recover the rotation periods of \naigrain\ simulated
\kepler-like light curves for solid-body rotators \citep{Aigrain2015} using
three different methods: a new Gaussian process method,
an authocorrelation function method, and a Lomb-Scargle periodogram method.
We demonstrate that the GP method produces the most accurate rotation periods
of the three techniques, providing a large improvement over the LS periodogram
method and a moderate one over our basic implementation of the ACF method.
We also find that our basic ACF method which is similar to methods implemented
in the literature, often slightly under-predicts the rotation period due to a
subtle feature of its peak detection algorithm.
Comparing our results with the results of the \citet{Aigrain2015} Tel Aviv
team who implement the method of \citet{Mcquillan2013}, a slightly different
version of the ACF method to the one we use, we find that their results are
slightly more precise than the GP method results overall.
They do however have a larger number of significant outliers.
We measured the rotation periods of \nkoimcq\ \kepler\ objects of interest
using the GP method, and find that these results compare well to those
measured previously by \citet{Mcquillan2013}.
The good agreement between these two sets of results demonstrates that the GP
method works well on real \kepler\ data.
